---
title: "COVID-19 around the world"
author:
 - name: "Annette Lien (s202633)"
 - name: "Carlota Carbajo Moral (s202424)"
 - name: "Mathias Fynbo Jensen (s192065)"
 - name: "Thomas Carvalho Andersen (s202755)"
date: "10/5/2021"
output: 
  ioslides_presentation:
    logo: presentation-figure/dtu-logo-png-transparent.png
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Index

- Main idea
- Data loading
- Data cleaning
- Data augmentation
- Data anlysis
- Conclusion

## Main idea

For this project we merged data from two main sources: John Hopkins and Gapminder.

<div class="columns-2">
### John Hopkins data - [Repository Link](https://github.com/CSSEGISandData/COVID-19)

- Main datasets of the project
  - Time series of COVID-19 confirmed cases
  - Time series of COVID-19 deaths
  - Population count (IUD)
  


### Gapminder data - [Website Link](https://www.gapminder.org/data/)

- 18 datasets of varying relevance including:  
  - Life expectancy
  - Smoking and Alcohol Consumption
  - Body Mass Index and Basic Sanitation

</div>

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

